### 2019-02-15 paper/birth_spacing_lit.tex
Update lit notes

@@

@@

several papers show empirically that in settings with son preference, a
couple that just had a son is more likely to stop having children (Das
1987) or wait longer to have the next child (Rindfuss et al. 1982;
Trussell et al. 1985; Arnold, Choe, and Roy 1998; RetherfordandRoy
2003). Hemochandra, Singh, and Singh (2010) show that birth intervals
are shorter if parents have not reached their self-reported ideal number
of sons. Second, Nemeth and Bowling (1985) examine whether having nosons
shortens the duration of breastfeeding.

@@

We ﬁnd that the estimates of birth spacing on child mortality are different when we do not account for fertility selection.

@@

India: There is evidence of a statistically signiﬁcant effect of birth spacing on child survival: an increase in the duration between child i and child i + 1 signiﬁcantly reduces the hazard of mortality of child i (equivalently increases the survival chances of the child).
for the sample of Pakistani households. As with the Indian case, there is evidence of maternal depletion/sibling competition effect in the Pakistani sample: an increase in the duration between child i and child i + 1 signiﬁcantly reduces the hazard of mortality of child i.

@@

This article assesses the strength of son preference in Vietnam, as reflected infertility behavior. It
formulates and estimates a proportional hazards model applied to birth intervals, and a contraceptive
prevalence model, using household survey datafrom 2,636 ever-married women aged 15-49 with at
least one living child who were interviewedfor the Vietnam Living Standards Survey 1992-1993.

>>>>

we use dynamic panel data techniques to analyse the causal effects of infant mortality on birth intervals and fertility, as well as the causal effects of birth intervals on mortality in rural Bangladesh, accounting for unobserved heterogeneity and reverse causality

@@

We find evidence of causal effects in two directions: a short preceding birth interval reduces survival chances of infants, and an infant death increases the probability of a next birth and shortens the time until the next birth (replacement behaviour)

Estimates of reproductive behaviour are consistent with gender preference: having more surviving boys significantly reduces the probability of having a next child and this effect is strongest in the comparison area

@@

several papers show empirically that in settings with son preference, a
couple that just had a son is more likely to stop having children (Das
1987) or wait longer to have the next child (Rindfuss et al. 1982;
Trussell et al. 1985; Arnold, Choe, and Roy 1998; RetherfordandRoy
2003). Hemochandra, Singh, and Singh (2010) show that birth intervals
are shorter if parents have not reached their self-reported ideal number
of sons. Second, Nemeth and Bowling (1985) examine whether having no sons
shortens the duration of breastfeeding.

@@

We find that the estimates of birth spacing on child mortality are
different when we do not account for fertility selection.

@@

India: There is evidence of a statistically signifiant effect of birth
spacing on child survival: an increase in the duration between child i
and child i + 1 significantly reduces the hazard of mortality of child i
(equivalently increases the survival chances of the child).
for the sample of Pakistani households. As with the Indian case, there
is evidence of maternal depletion/sibling competition effect in the
Pakistani sample: an increase in the duration between child i and child
i + 1 significantly reduces the hazard of mortality of child i.

@@

This article assesses the strength of son preference in Vietnam, as reflected in fertility behavior. It
formulates and estimates a proportional hazards model applied to birth intervals, and a contraceptive
prevalence model, using household survey data from 2,636 ever-married women aged 15-49 with at
least one living child who were interviewedfor the Vietnam Living Standards Survey 1992-1993.
